---
title: "Who Joined Code for Philly Slack Channels in the Last 10 Days?"
author: Michael Chow
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(dbplyr)

# install.packages("slackr")
library(slackr)

# note that I had to patch slackr to pull more than 1000 users ----
source("fix-slackr.R", local = M <- new.env())

# utility funcs ----
frame_to_listcol <- function(df_col) {
  # data.frames can have columns that are data.frames. It's a bad
  # idea, so we turn them into list columns.
  split(df_col, seq(nrow(df_col)))  
}

# make sure we have a token ----
stopifnot(
  Sys.getenv("SLACK_API_TOKEN") != ""
  )
```

## Fetching data

Uses the following slackr functions:

* slackr_users : get a data frame of Slack
* slackr_channels : get a data frame of Slack
* slackr_groups : get a data frame of Slack groups
* slackr_history: Return message history of a Slack channel to a data.frame

```{r fetch_raw, cache = TRUE}
#START_DATE <- now() - ddays(10)
#END_DATE <- now()

raw_users <-
  #slackr_users() %>%
  M$slackr_fixed_users() %>%
  as_tibble()# %>%
  #select(id, user_name = name)

raw_channels <-
  slackr_channels() %>%
  filter(!is_archived)

fetch_history <- function(channel) {
  # slack api rate limits at 50 / minute for this kind of request
  res <- history_slackr(1000, channel = paste0('#', channel))
  # slackr returns inconsistent types, so we use only tibbles... :/
  if (is_empty(res)) return(tibble())
  res
  #select(res, user, text, ts)
}

raw_messages_nested <-
  raw_channels %>%
  select(channel_name = name) %>%
  mutate(
    history = map(
      channel_name,
      slowly(~ fetch_history(.x),  rate_delay(1.25))
    )
  )


# sanity check that dataframe columns are well-structured
# they should have same # of rows as data
# this is kind of nuts, but that's what happens when columns are data.frames
fix_raw <-
  raw_messages_nested %>%
  rowwise() %>%
  do(
    valid_subdfs = all(
      map_int(keep(.$history, ~ "data.frame" %in% class(.x)), nrow) == nrow(.$history)
    ),
    fixed = mutate_if(.$history, ~ is(., "data.frame"), frame_to_listcol),
    channel_name = .$channel_name
  ) %>%
  mutate(
    valid_subdfs = as.logical(valid_subdfs),
    channel_name = as.character(channel_name)
    )
  #{stopifnot(all(.$valid_subdfs))}
  
  
raw_messages <-
  fix_raw %>%
  unnest(fixed) %>%
  mutate(ts = as_datetime(as.numeric(ts)))
```

## Wrangling

Here, I just removed all the columns holding nested data. Would be quick to add useful ones as tables!

```{r wrangle}
channels <-
  raw_channels %>%
  select(-topic, -purpose, -previous_names, -members) %>%
  as_tibble()

messages <-
  raw_messages %>%
  select(
    channel_name, type, subtype, text, ts, username, bot_id, user, team, client_msg_id,
    thread_ts, reply_count, reply_users_count, latest_reply, subscribed,
    parent_user_id, last_read, is_starred, item_type, upload_reply_to, topic,
    purpose, inviter, old_name, name, is_intro, hidden, bot_link, new_broadcast,
    channel, no_notifications, permalink
  )

users <-
  raw_users %>%
  select(id, real_name, display_name, title, team_id, deleted, updated, status_text, is_invited_user, bot_id, team, guest_invited_by)
```

## Create DB Tables

```{r put-in-db}
con <- DBI::dbConnect(RPostgres::Postgres(), 
  host = "localhost",
  port = 5434,
  user = "postgres",
  password = "",
  dbname = "postgres"
)
```

```{r}
copy_slack <- function(con, df) {
  prefixed <- paste("slack", deparse(substitute(df)), sep = "_")
  copy_to(con, df, prefixed, temporary = FALSE, overwrite = TRUE)
}

copy_slack(con, channels)     # creates slack_channels table
copy_slack(con, messages)
copy_slack(con, users)
```

```{r, eval=FALSE}
write_csv(channels, "channels.csv")
write_csv(messages, "messages.csv")
write_csv(users, "users.csv")
```


## Demo of querying DB

```{r}
# example counting number of archived channels using db
tbl(con, "channels") %>% count(is_archived)
```

```{r}
tbl(con, "messages") %>% group_by(channel_name)
```



